---
title: "FML Assignment4"
author: "Priyanka Jonnala"
date: "2023-11-12"
output:
  word_document: default
  pdf_document: default
  html_document: default
---
Loading the required packages
```{r}
library(flexclust)
library(cluster)
library(ggcorrplot)
```
Importing the dataset
```{r}
Pharmaceuticals <- read.csv("C:\\Users\\priya\\OneDrive\\Desktop\\FML\\Assignment 4\\Pharmaceuticals.csv")
head(Pharmaceuticals)
```
Choose columns 3 to 11 and store the resulting data frame in Pharma1
```{r}
Pharma <- na.omit(Pharmaceuticals)
Pharma
Pharma1 <- Pharma[,3:11]
# Displaying the top six rows of Pharma1 using head function
head(Pharma1)
# Printing summary statistics for Pharma1
summary(Pharma1)
```

```{r}
library(factoextra)
```
Normalizing the data with scale method
```{r}
Pharma2 <- scale(Pharma1)
# Set row names to match the first column of the original Pharma data
row.names(Pharma2) <- Pharma[,1]
# Calculate the distance matrix using get_dist
distance <- get_dist(Pharma2)
# Visualize the distance matrix using fviz_dist
fviz_dist(distance)
```

```{r}
# Set the random seed for reproducibility
set.seed(10)
# Use a for loop to calculate the within-cluster sum of squares (wss) for 1 to 10 clusters
wss <- vector()
for(i in 1:10) wss[i] <- sum(kmeans(Pharma2,i)$withinss)
# Visualize the wss values using a line plot
plot(1:10, wss , type = "b" , main = paste('Cluster of Companies') , xlab =
"Number of Clusters", ylab="wss")
# Print the wss values for each number of clusters
wss
```
Silhouette Approach
```{r}
# Use the fviz_nbclust function to determine the optimal number of clusters using the silhouette method
fviz_nbclust(Pharma2, kmeans, method = "silhouette")
```
This demonstrates that five clusters are the optimum number. Using the k-means method to create a 5 cluster.
```{r}
# Use the kmeans function to create 5 clusters and visualize the results using the fviz_cluster function
set.seed(21)
clusterx <- kmeans(Pharma2, centers = 5, nstart = 25) # k = 5, number of restarts = 25
clusterx$centers
clusterx$size
fviz_cluster(clusterx, data = Pharma2)
```
Manhattan Distance when using Kmeans Clustering.
```{r}
set.seed(21)
# Use kcca function to create 5 clusters with Manhattan distance and k-medians algorithm
clusterY <- kcca(Pharma2, k = 5, kccaFamily("kmedians"))
# Print the results and visualize the clusters
clusterY
clusters_index <- predict(clusterY)
dist(clusterY@centers)
image(clusterY)
points(Pharma2, col = clusters_index, pch = 19, cex = 0.3)
```

```{r}
library(tidyverse)
```

```{r}
Pharma1 %>% mutate(Cluster = clusterx$cluster) %>% group_by(Cluster) %>% summarise_all("mean")
clusplot(Pharma2,clusterx$cluster, main="Clusters",color = TRUE,shade = TRUE, labels = 2,lines = 0)
```
Companies are classified into different clusters as follows

Cluster1 : AHM,WYE,BMY,AZN,LLY,ABT,NVS,ABT and SGP

Cluster2 : AGN,PHA

Cluster3 : BAY,CHTT,IVX

Cluster4 : ELN,MRX,WPI,AVE

Cluster5 : JNJ,MRK,PFE,GSK

From the means of the cluster variables, it can be derived as follow:

Cluster1 has a medium risk

Cluster2 has very high PE Ratio

Cluster3 Despite having an excellent PE ratio, it is incredibly risky to own due to its extremely high risk, extremely high leverage, and poor Net Profit margin. Also very low in revenue growth.

Cluster4 has the best Net Profit Margin, the lowest PE ratio, and the fastest sales growth. It can be bought or kept on hand as a reserve.

Cluster5 has strong market capitalization, ROI, ROA, ROA on assets, ROA on turnover of assets, and ROA on net profit margin. A low PE ratio indicates that the stock price is moderately valued and may thus be bought and kept. Revenue growth of 18.5% is also favorable.

Examining patterns by visualizing clusters against the variables
```{r}
Pharma3 <- Pharma[12:14] %>% mutate(Clusters=clusterx$cluster)
ggplot(Pharma3, mapping = aes(factor(Clusters), fill =Median_Recommendation)) +
 geom_bar(position='dodge') + labs(x ='Clusters')
```

```{r}
ggplot(Pharma3, mapping = aes(factor(Clusters),fill = Location)) +
 geom_bar(position = 'dodge') + labs(x ='Clusters')
```

```{r}
ggplot(Pharma3, mapping = aes(factor(Clusters),fill = Exchange)) +
 geom_bar(position = 'dodge') + labs(x ='Clusters')
```

Provide an appropriate name for each cluster using any or all of the variables in the data set.

Cluster1: Attempt it Cluster 

Cluster2: Significant Risk Cluster

Cluster3: Very Dangerous Cluster

Cluster4: Top Buying Cluster

Cluster5: A Perfect Asset